---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---

AUTHOR: Greeshma Chowdary Peddineni
```{r warning=FALSE, message=FALSE}
if(!require(EDAWR)) {
install.packages("devtools")
devtools::install_github("rstudio/EDAWR")
}
library(tidyr)
library(dplyr)
library(EDAWR)
library(ggplot2)

```

Using the STORMS dataset, and DPLYR, answer the following:
Your answer should display directly in the knitted document and include only the answer (no extraneous info)
1:
show only storm (name) and date
```{r}

storms %>% mutate(date = paste(year, month, day, sep = "-")) %>% select(name, date)
```

2:
show only storms where wind was less than 55
```{r}

filter(storms, wind<55)
```

3: 
split date into individual columns for month, day, year
save as a new dataset called mystorms
```{r}

mystorms <- storms %>% 
  unite("date", year:day, sep = "-") %>% 
  separate(date, into = c("year", "month", "day"), sep="-")
mystorms
```

4:
list storm name(s) for any storms that happened in June (from your mystorms dataset)
```{r}

mystorms %>% filter(month == 6) %>% select(name)
```

Now using the rates dataset, answer the following:

5:
what is the rate for Ethiopia in 2007?
```{r}

rates %>% filter(country == "Ethiopia" & year == 2007) %>% select(rate)
```

6:
create a copy of the rates dataset called myrates
add a new column that is population times rate, call this column 'cases'
which country has the highest value in cases, in which year?
```{r}
myrates <- rates %>% mutate(cases = population * rate)
highest_cases <- myrates %>% arrange(desc(cases)) %>% slice(1)
highest_cases
```

7:
Change year to a factor in myrates.
create a plot for Ethiopia showing cases over time
```{r}
ethiopia_cases <- myrates %>% filter(country == "Ethiopia")
ggplot(ethiopia_cases, aes(x = year, y = cases)) + geom_col(fill = "maroon") + labs(title = "Ethiopia: Cases Over Time", x="Year", y = "Cases") +
theme_minimal()
myrates <- myrates %>% mutate(cases_per_1000 = cases / 1000)
myrates

```


